
# R Programming - Associative Analysis



```{r}
# We first we install the required arules library 

install.packages("arules")
```


```{r}
# Loading the arules library
#
library(arules)
```




```{r}
# Loading our transactions dataset from our csv file

Transactionss <- read.transactions("http://bit.ly/SupermarketDatasetII")
View(Transactionss)
```


```{r}
# Verifying the object's class type of data

class(Transactionss)
```


```{r}

# Previewing our first 5 transactions

inspect(Transactionss[1:5])
```


```{r}
# Previewing the items that make up our dataset

items<-as.data.frame(itemLabels(Transactionss))
colnames(items) <- "Item"
head(items, 10)    
```


```{r}
# Getting summary of the transaction dataset

summary(Transactionss)
```


```{r}
# Exploring the frequency of some parchased items

itemFrequency(Transactionss[, 8:10],type = "absolute")
round(itemFrequency(Transactionss[, 8:10],type = "relative")*100,2)
```


```{r}
# Producing a chart of frequencies and fitering  items that has minimum percentage first ten
# to consider only items with a minimum percentage 
 
par(mfrow = c(1, 2))

# plot the frequency of items
itemFrequencyPlot(Transactionss, topN = 10,col="purple")
itemFrequencyPlot(Transactionss, support = 0.1,col="darkblue")
```
from the graph above we can observe that tea has the highest frequency meaning it has a high sales with grated french being the lowest sales.

```{r}
#  Using apriori will building a model based on association rules 
# With confidence as 0.8  and min Support as 0.001 
 
rules <- apriori (Transactionss, parameter = list(supp = 0.001, conf = 0.8))
rules

```


```{r}
# We use measures of significance and interest on the rules, 
# determining which ones are interesting and which to discard.
# we will see what happens if we increase the support or lower the confidence level
# 

# Using apriori will building a model based on association rules  with Min Support as 0.002 and confidence as 0.8.
rules2 <- apriori (Transactionss,parameter = list(supp = 0.002, conf = 0.8)) 

# Using apriori will building a model based on association rules  with Min Support as 0.002 and confidence as 0.6.
rules3 <- apriori (Transactionss, parameter = list(supp = 0.001, conf = 0.6)) 

rules2

rules3
```
In 15 counts and 7 counts we see that their is an increase from 99 to 319



```{r}
# Check the model summary

summary(rules)
```


```{r}
# Observing first 5 model rules  built in our model 
# ---
# 
inspect(rules[1:5])

```


```{r}
# Ordering these rules by a criteria such as the level of confidence
# then looking at the first five rules.
# We can also use different criteria such as: (by = "lift" or by = "support")
# 
rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:5])

```


```{r}
# If we're interested in making a promotion relating to the sale of yogurt, 
# we could create a subset of rules concerning these products 
# ---
# This would tell us the items that the customers bought before purchasing yogurt
# ---
# 
wheat <- subset(rules, subset = rhs %pin% "wheat")
 
# Then order by confidence
wheat<-sort(wheat, by="confidence", decreasing=TRUE)
inspect(wheat[1:5])
```


```{r}
# What if we wanted to determine items that customers might buy 
# who have previously bought yogurt?
# ---
# 
# Subset the rules
wheat <- subset(rules, subset = lhs %pin% "wheat")

# Order by confidence
wheat<-sort(wheat, by="confidence", decreasing=TRUE)

# inspect top 5
inspect(wheat[5:10])
```
the chances of one buying wheat to nine has a confidence of of 100% 
In recommendation the first nine products should be a priority and sealer should priorities on those products.
